Distributional Trees and Forests

### 2017-12-08 pkg/disttree/inst/CMStatistics2017/slides_rain.Rnw
resized map and small corrections
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1828 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: ea82a0595ee89a9a4db0f8fa815d3dfc0f249111

\documentclass[11pt,t,usepdftitle=false,aspectratio=43]{beamer}
\usetheme[nototalframenumber, noslidenumber, url]{uibk}

@@

\textbf{Idea:} Statistical model with numerical predictions as covariates.\\
\bigskip
\textbf{Distribution assumption:}\\
\vspace{-0.3cm}
%\hspace{1.5cm}
%\center
$$(\text{precipitation})^\frac{1}{1.6} \sim \textit{c}\mathcal{N}(\mu,\sigma^2) \quad \text{censored at 0.}$$

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.2cm}
\textbf{Application for one station:}
\begin{itemize}
\item Learn forest model on data from 24 years.
\item Evaluate on 4 years.
\item 10 times 7-fold cross validation.
\item Compare to commonly used heteroscedastic censored Gaussian models:
\end{itemize}
%\begin{itemize}
%\item Prespecified GAM: variable selection based on meteorological expert knowledge.
%\item Boosted GAM: automatic variable selection.
%\item Ensemble Model Output Statistics (EMOS): linear predictors using only total precipitation predictions as covariates.
%\end{itemize}
\vspace{0.1cm}
\begin{table}[t!]
\footnotesize
\nodindent
\hskip-0.4cm\begin{tabular}{ l l l }
\hline
Model & Type & Variable Selection \\
\hline
\vspace{-0.15cm}
\textbf{Forest} & recursive    & automatic \\ 
                & partitioning &           \\ 
\hline
\vspace{-0.15cm}
\textbf{prespecified GAM} & spline  & based on expert \\
                          & in each & knowledge      \\
\hline
\vspace{-0.15cm}
\textbf{boosted GAM} & spline  & automatic  \\
                     & in each &            \\
\hline
\vspace{-0.15cm}
\textbf{EMOS} & linear & total precipitation mean \\
              &        & and standard deviation   \\
\hline
\end{tabular}
\end{table}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\begin{center}
\vspace*{0cm}

<<map, fig=TRUE, echo=FALSE, width=10, height = 6.5>>=

layout(cbind(1, 2), width = c(9, 1))
par(mar = c(5,4,4,0.1))
raster::image(tirol.dem, col = rev(gray.colors(100)),
              main="Observation stations", ylab = "Latitude", xlab = "Longitude", 
              xlim = c(9.8,13.2), 
              ylim = c(46.6, 47.8), 
              cex.lab=1.5, cex.axis=1.2, cex.main=2)
plot(tirol.gadm, add = TRUE)
points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.9)
legend("topleft", pch=c(21,24,25,22), legend = c("Forest", "prespecified GAM", "boosted GAM", "EMOS"), cex = 1.2, bty = "n")

par(mar = c(0.5,0.2,0.5,2.3))
## legend
plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
     xlim = c(0, 1), ylim = c(-0.2, 0.2), xaxs = "i", yaxs = "i")
rect(0, brk[-6], 0.5, brk[-1], col = rev(clr))
axis(4, at = brk, las = 1, mgp=c(0,0,-1.2), cex = 1.4)

>>>>

\documentclass[11pt,t,usepdftitle=false,aspectratio=43]{beamer}
\usetheme[nototalframenumber, noslidenumber]{uibk}

@@

\textbf{Idea:} Statistical model with numerical predictions as covariates.\\
\bigskip
\textbf{Distribution assumption:}\\
\vspace{-0.25cm}
%\hspace{1.5cm}
%\center
$$(\text{precipitation})^\frac{1}{1.6} \sim \textit{c}\mathcal{N}(\mu,\sigma^2) \quad \text{censored at 0.}$$

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.2cm}
\textbf{Application for one station:}
\begin{itemize}
\item Learn forest model on data from 24 years.
\item Evaluate on 4 years.
\item 10 times 7-fold cross validation.
\item Compare to commonly used heteroscedastic censored Gaussian models:
\end{itemize}
%\begin{itemize}
%\item Prespecified GAM: variable selection based on meteorological expert knowledge.
%\item Boosted GAM: automatic variable selection.
%\item Ensemble Model Output Statistics (EMOS): linear predictors using only total precipitation predictions as covariates.
%\end{itemize}
\vspace{0.1cm}
\begin{table}[t!]
\footnotesize

\hskip-0.4cm\begin{tabular}{ l l l }
\hline
Model & Type & Variable Selection \\
\hline
\vspace{-0.15cm}
\textbf{Forest} & recursive    & automatic \\ 
                & partitioning &           \\ 
\hline
\vspace{-0.15cm}
\textbf{prespecified GAM} & spline  & based on expert \\
                          & in each & knowledge      \\
\hline
\vspace{-0.15cm}
\textbf{boosted GAM} & spline  & automatic  \\
                     & in each &            \\
\hline
\vspace{-0.15cm}
\textbf{EMOS} & linear & total precipitation mean \\
              &        & and standard deviation   \\
\hline
\end{tabular}
\end{table}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\begin{center}
%\vspace*{-2cm}

\setkeys{Gin}{width=0.95\textwidth}
<<map, fig=TRUE, echo=FALSE, width=10, height=6.5>>=

layout(cbind(1, 2), width = c(9, 1))
par(mar = c(5,4,4,0.01))
raster::image(tirol.dem, col = rev(gray.colors(100)),
              main="Observation stations", ylab = "Latitude", xlab = "Longitude", 
              xlim = c(9.8,13.2), 
              ylim = c(46.6, 47.85), 
              cex.lab=1.3, cex.axis=1.2, cex.main=1.7)
plot(tirol.gadm, add = TRUE)
points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.7)
legend("topleft", pch=c(21,24,25,22), legend = c("Forest", "prespecified GAM", "boosted GAM", "EMOS"), cex = 1.2, bty = "n")

par(mar = c(0.4,0.2,0.4,2.33))
## legend
plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
     xlim = c(0, 1), ylim = c(-0.2, 0.2), xaxs = "i", yaxs = "i")
rect(0, brk[-6], 0.5, brk[-1], col = rev(clr))
axis(4, at = brk, las = 1, mgp=c(0,0,-1.2), cex = 1.5)
